---
title: "cbsodataR, all data of Statistics Netherlands (CBS)"
author: "Edwin de Jonge"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{cbsodataR, all data of Statistics Netherlands (CBS)}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include=FALSE}
library(dplyr)
```

[Statistics Netherlands (CBS)](http://www.cbs.nl) is the office that produces all official statistics of the Netherlands. 

For long SN has put its data on the web in its online database 
[StatLine](http://statline.cbs.nl/Statweb/?LA=en).
Since 2014 this data base has an open data web API based on the OData protocol.
The *cbsodataR* package allows for retrieving data right into R. 

## Table of Contents

A list of tables can be retrieved using the `cbs_get_toc` function.

```{r cbs_get_toc, message=FALSE}
library(dplyr) # not needed, but used in examples below
library(cbsodataR)

toc <- cbs_get_toc(Language="en") # retrieve only english tables

toc %>% 
  select(Identifier, ShortTitle) 
```

Using an "Identifier" from `cbs_get_toc` information on the table can be retrieved with `cbs_get_meta` 

```{r get_meta, message=FALSE}
apples <- cbs_get_meta('71509ENG')
apples
```
The meta object contains all metadata properties of cbsodata (see the [original documentation](http://www.cbs.nl/NR/rdonlyres/2561A2B7-CE51-47B9-A838-9968EF67FFB0/0/2014handleidingcbsopendataservices.pdf)) in the form of data.frames. Each data.frame describes properties of the SN table.

```{r, meta2}
names(apples)
```

## Data download

With `cbs_get_data` data can be retrieved. By default all data for this table will be downloaded in a temporary directory.
```{r, cbs_get_data2, message=FALSE}
cbs_get_data('71509ENG') %>% 
  select(1:4) %>%  # demonstration purpose
  head()
```

### Adding category labels

The first columns are categorical columns: they contain codes. The labels for these columns can be added with
the function `cbs_add_label_columns`. 

```{r, get_data, message=FALSE}
cbs_get_data('71509ENG') %>%
  cbs_add_label_columns() %>% 
  select(1:4) %>% 
  head()
```

### Adding Date column

The period/time columns of Statistics Netherlands (CBS) contain coded time periods:
e.g. 2018JJ00 (i.e. 2018), 2018KW03 (i.e. 2018 Q3), 2016MM04 (i.e. 2016 April).
With `cbs_add_date_column` the time periods will be converted and added to the data:

```{r}
cbs_get_data('71509ENG') %>%
  cbs_add_date_column() %>% 
  select(2:4) %>% 
  head()
```

This can be useful for further time series analysis, but also for displaying.
It is also possible to convert the dates to numbers:
```{r}
cbs_get_data('71509ENG') %>%
  cbs_add_date_column(date_type = "numeric") %>% 
  select(2:4) %>% 
  head()
```



## Select and filter

It is possible restrict the download using filter statements. This may shorten the download time considerably.

```{r, get_data3, message=FALSE}
  cbs_get_data('71509ENG', Periods='2000JJ00') %>% 
  select(1:4) %>% 
  head()
```


# Download data

Data can also be downloaded explicitly by using `cbs_download_table`

