---
title: "cbsodata, all data of Statistics Netherlands (CBS)"
author: "Edwin de Jonge"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  \usepackage[utf8]{inputenc}
---
```{r init, echo=FALSE, message=FALSE}
library(dplyr, quietly = T)
library(cbsodata)
```


Statistics Netherlands (CBS) is the office that produces all official statistics of the Netherlands. 

For long SN has put its data online in its online database 
[StatLine](http://statline.cbs.nl/Statweb/?LA=en).
Since 2014 this data base has an open data web API based on the OData protocol.
The *cbsodata* package allows for retrieving data right into R. 

## Table information

A list of tables can be retrieved using the `get_table_list` function.

```{r get_table_list, message=FALSE}
tables <- get_table_list(Language="en") # retrieve only enlgish tables

tables %>% 
  select(Identifier, ShortTitle) %>% 
  head 
```

Using an "Identifier" from `get_table_list` information on the table can be retrieve with `get_meta` 

```{r get_meta, message=FALSE}
m <- get_meta('71509ENG')
m
```
The meta object contains all metadata properties of cbsodata (see the original documentation for that) in the form
of data.frames

```{r, meta2}
names(m)
```

## Data download

With `get_data` data can be retrieved. By default all data for this table will be downloaded in a temporary directory.
```{r, get_data2, message=FALSE}
get_data('71509ENG') %>% 
  select(2:5) %>% 
  head
```

The data will be automatically recoded with titles for the categories. If needed the original data can be retained with 
`recode=FALSE`

```{r, get_data, message=FALSE}
get_data('71509ENG', recode = FALSE) %>% 
  select(2:5) %>% 
  head
```

