---
title: 
author: 
date: 
output: 
  github_document: default
---

```{r, include=FALSE}
library(dplyr)
```

[Statistics Netherlands (CBS)](https://www.cbs.nl) is the office that produces all official statistics of the Netherlands. 

For long SN has put its data on the web in its online database 
[StatLine](https://opendata.cbs.nl/statline#/CBS/en/).
Since 2014 this data base has an open data web API based on the OData protocol.
The *cbsodataR* package allows for retrieving data right into R. 

## Table of Contents

A list of tables can be retrieved using the `cbs_get_toc` function.

```{r cbs_get_toc, message=FALSE}
library(dplyr) # not needed, but used in examples below
library(cbsodataR)

toc <- cbs_get_toc(Language="en") # retrieve only english tables

toc %>% 
  select(Identifier, ShortTitle) 
```

Using an "Identifier" from `cbs_get_toc` information on the table can be retrieved with `cbs_get_meta` 

```{r get_meta, message=FALSE}
apples <- cbs_get_meta('71509ENG')
apples
```
The meta object contains all metadata properties of cbsodata (see the [original documentation](https://www.cbs.nl/-/media/_pdf/2017/13/handleiding-cbs-open-data-services.pdf?la=nl-nl)) in the form of data.frames. Each data.frame describes properties of the SN table.

```{r, meta2}
names(apples)
```

## Data download

With `cbs_get_data` data can be retrieved. By default all data for this table will be downloaded in a temporary directory.
```{r, cbs_get_data2, message=FALSE}
cbs_get_data('71509ENG') %>% 
  select(1:4) %>%  # demonstration purpose
  head()
```
## Data download from a link

The opendata portal of CBS (https://opendata.cbs.nl/dataportaal/#/CBS/en) allows for finding a table and making
a selection within this table manually. Such a selection can be stored in a hyperlink (click the "share" button).
This link can also be used with `cbsodataR` using the `cbs_get_data_from_link` function. 

```{r}
cbs_get_data_from_link("https://opendata.cbs.nl/dataportaal/#/CBS/en/dataset/71509ENG/table?dl=193CB") %>% 
  select(1:4) %>% 
  head()
```


### Adding category labels

The first columns are categorical columns: they contain codes. The labels for these columns can be added with
the function `cbs_add_label_columns`. 

```{r, get_data, message=FALSE}
cbs_get_data('71509ENG') %>%
  cbs_add_label_columns() %>% 
  select(1:4) %>% 
  head()
```

### Adding Date column

The period/time columns of Statistics Netherlands (CBS) contain coded time periods:
e.g. 2018JJ00 (i.e. 2018), 2018KW03 (i.e. 2018 Q3), 2016MM04 (i.e. 2016 April).
With `cbs_add_date_column` the time periods will be converted and added to the data:

```{r}
cbs_get_data('71509ENG') %>%
  cbs_add_date_column() %>% 
  select(2:4) %>% 
  head()
```

This can be useful for further time series analysis, but also for displaying.
It is also possible to convert the dates to numbers:
```{r}
cbs_get_data('71509ENG') %>%
  cbs_add_date_column(date_type = "numeric") %>% 
  select(2:4) %>% 
  head()
```



## Select and filter`

It is possible restrict the download using filter statements. This may shorten the download time considerably.

### Filter

Filter statements for the columns can be used to restrict the download. Note the following:

- To filter you will need to use the values found in the `$Key` column in the `cbs_get_meta` objects. e.g. for year 2020, the code is "2020JJ00".

```{r}
apples <- cbs_get_meta('71509ENG')
names(apples)
# meta data for column Periods
head(apples$Periods[,1:2])
#meta data for column FruitFarmingRegions
head(apples$FruitFarmingRegions[,1:2 ])
```


- To filter for values in a column add `<column_name> = values` to `cbs_get_data` e.g. `Periods = c("2019JJ00", "2020JJ0")`


```{r, get_data3, message=FALSE}
  cbs_get_data( '71509ENG'
              , Periods=c('2000JJ00','2001JJ00') # selection on Periods column
              , FruitFarmingRegions = "1" # selection on FruitFarmingRegions
              #
              # restrict the columns to the following as found in
              # apples$DataProperties with "select"
              , select = c("FruitFarmingRegions", "Periods", "TotalAppleVarieties_1")  
              ) %>% 
  cbs_add_label_columns()
```

- To filter for values in a column that have a substring e.g. "JJ" you can use `<column_name> = has_substring(<substring>)` to `cbs_get_data` e.g. `Periods = has_substring("KW")`

```{r}
  cbs_get_data( '71509ENG'
              , Periods = has_substring('2000') # selection on Periods column
              , FruitFarmingRegions = "1" # selection on FruitFarmingRegions
              #
              # restrict the columns to the following as found in
              # cbs_get_meta("71509ENG")$DataProperties with "select"
              , select = c("FruitFarmingRegions", "Periods", "TotalAppleVarieties_1")  
              ) %>% 
    cbs_add_label_columns()

```

- To combine values and substring use the "|" operator: `Periods = eq("2020JJ00") | has_substring("KW")`

```{r}
  cbs_get_data( '71509ENG'
              , Periods = eq("2010JJ00") | has_substring('2000') # selection on Periods column
              , FruitFarmingRegions = "1" # selection on FruitFarmingRegions
              #
              # restrict the columns to the following as found in
              # cbs_get_meta("71509ENG")$DataProperties with "select"
              , select = c("FruitFarmingRegions", "Periods", "TotalAppleVarieties_1")  
              ) %>% 
    cbs_add_label_columns()
```

# Download data

Data can also be downloaded explicitly by using `cbs_download_table`

